Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-07 paper/paper_new.tex
Updates from Overleaf

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results and Discussion}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 50 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

@@

Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box.
50 randomly-chosen images were reviewed manually to evaluate the
calculation, a representative sample are presented in
Figure \ref{fig:bbox_mask}.
All 50 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 22 masks and bounding boxes covered the entire fish,
16 missed some of the tail, and 12 missed most or all of the tail, as
seen in Figure \ref{fig:missedTail}.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/61631}
  \includegraphics[width=0.49\linewidth]{images/61631_mask}
  \caption{An example of a light colored tail being missed during the pixel analysis process.}
  \label{fig:missedTail}
\end{figure}
Masks and bounding boxes contain the head and trunk of the fish in nearly all cases, but further refinement of our algorithms will be needed
to ensure that light fins and tails are masked consistently and accurately.

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 50 test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.91\%\), and the average error for the fish length calculation was \(5.87\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}

@@

\section{Conclusions}
The BGNN project demonstrates a clear need for improved metadata for specimen image collections. Much effort, time and money have been put into photographing and digitizing physical specimen, but without detailed and complete metadata properties the utility of these repositories for advanced computation analysis and machine learning is limited. Since it is prohibitively time intensive to generate all the pertinent properties by hand, automated techniques are needed to generate this missing metadata at scale.

In this paper we presented such a metadata generation program. It produces 6 of the 22 core BGNN metadata properties~\cite{leipzig2021biodiversity}, as well as image contrast, bounding boxes and scale and length information. Testing this program on \(7,244\) images from the INHS dataset~\cite{INHS}, we see that the vast majority of the resulting metadata is correct within a tolerance of a few percentage points, and in some cases contains fewer mistakes than the manually generated validation metadata. Through further refinement and generalizing beyond only INHS images, we aim to create a tool that can be distributed to specimen image collection curators to correct the metadata sparsity that precipitated this work.
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may be able to improve masking of such features. Another possible approach to solving this is to threshold and mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process to work on classes of specimen images (specifically all fish images for BGNN), not only the INHS collection. To accomplish this, a much larger training dataset full of more diverse images will be required. Another requirement will be to generalize the ruler reading process beyond the INHS specific reading of digits on the ruler, which will likely involve an automated method of reading ruler ticks instead of digits.
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}

>>>>

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results and Discussion}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 100 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

@@

Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box.
100 randomly-chosen images were reviewed manually to evaluate the
calculation, a representative sample are presented in
Figure \ref{fig:bbox_mask}.
All 100 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 41 masks and bounding boxes covered the entire fish,
36 missed some of the tail, and 23 missed most or all of the tail, as
seen in Figure \ref{fig:missedTail}.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/61631}
  \includegraphics[width=0.49\linewidth]{images/61631_mask}
  \caption{An example of a light colored tail being missed during the pixel analysis process.}
  \label{fig:missedTail}
\end{figure}
Masks and bounding boxes contain the head and trunk of the fish in nearly all cases, but further refinement of our algorithms will be needed
to ensure that light fins and tails are masked consistently and accurately.

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 100  test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.91\%\), and the average error for the fish length calculation was \(5.87\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}

@@

\section{Conclusions}
The BGNN project demonstrates a clear need for improved metadata associated with specimen image collections. Much effort, time and money have been put into photographing and digitizing physical specimen, but without detailed and complete metadata properties the utility of these repositories for advanced computation analysis and machine learning is limited. Since it is prohibitively time intensive to generate all the pertinent properties by hand, automated techniques are needed to generate this missing metadata at scale.

In this paper we presented such a metadata generation program. Using machine learning and image informatics tools and techniques, it is able to locate, bitmask and analyze specimen (currently limited to fish) in collection images with a high degree of accuracy. It produces 6 of the 22 core BGNN metadata properties~\cite{leipzig2021biodiversity}, as well as image contrast, bounding boxes and scale and length information. Testing this program on \(7,244\) images from the INHS dataset~\cite{INHS}, we see that the vast majority of the resulting metadata is correct within a tolerance of a few percentage points, and in some cases contains fewer mistakes than the manually generated validation data. Through further refinement and generalizing beyond only INHS images, we aim to create a tool that can be distributed to specimen image collection curators to correct the metadata sparsity that precipitated this work.
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may be able to improve masking of such features. Another possible approach to solving this is to threshold and mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process that works on classes of specimen images. For the BGNN project we are beginning with fish images, but we are designing the metadata generation system such that it can eventually operate on other species if trained that way. To accomplish this, a much larger training dataset full of more diverse images will be required. The first step towards greater generality will be to operate on other fish collections besides INHS, which is something our program has already show itself capable of doing during initial testing. Another requirement will be to generalize the ruler reading process beyond the INHS specific reading of digits on the ruler, which will likely involve an automated method of reading ruler ticks instead of digits.
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}
